---
output: github_document
---

<!-- Edit this to prevent git error: 9827634 -->
<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
# Set all better options to NULL
for (i in stringr::str_subset(names(options()),"better\\.")) {
  print(i)
  print(.Options[[i]])
  .Options[[i]] <- NULL
  print(.Options[[i]])
}
options(
  better.gbd_path = NULL #"/Users/jwl38/Library/CloudStorage/GoogleDrive-janwillem.lindemans@gmail.com/My\ Drive/Offline\ Drive/R\ -\ Offline\ Drive/R\ packages\ JW/better/ignore/Global\ Burden\ of\ Disease\ Study\ Data"
  )
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# better

IMPORTANT: THIS PACKAGE IS UNDER DEVELOPMENT. IT'S AVAILABLE HERE FOR DEVELOPMENT AND TESTING PURPOSES ONLY.

<!-- COPY from better.rmd -->

The `better` package helps you guesstimate the effect of health nudges, that is, behavioral interventions targeted at health behaviors. With `better` guesstimates, you can make the world a `better` place! In addition to guesstimating effect sizes, `better` also helps you guesstimate sample sizes for properly powered studies on health nudges. Finally, `better` also gives guesstimates of how much money you would save in terms of health care.

## Installation

You can install the development version of better from [GitHub](https://github.com/) with:

``` r
# install.packages("devtools")
devtools::install_github("janlindemans/better", build_vignettes = TRUE)
```

## Usage

<!-- COPY from better vignette: edit original there -->


## An example: A diet-related nudge for diabetics

<!-- ORIGINAL HERE - COPY PASTE it to elsewhere -->

```{r, include = FALSE}
library(better)
library(tidyverse)
gbd_filter()
#stop("gbd rows: ",nrow(gbd))
# So that the once-per-session messages are gone:
nudge("information") %>% 
  disease("diabetes mellitus")
```

The main purpose of `better` is to help you better guesstimate the effect of health nudges, that is, behavioral interventions targeted at health behaviors. 

Imagine you are designing a public health campaign. You are thinking about focusing on diabetics, and you are considering making information about healthy foods more accessible. What effects can you expect from a campaign like that? 

This is how we would make a guesstimate with `better`. Let's load it, and let's also load `tidyverse`, so we can use pipes and other tidy constructions.

```{r}
library(better)
library(tidyverse)

nudge("information", behavior = "food") %>% 
  disease("diabetes mellitus") %>%
  describe
```

Note that "DALYs" are *disability-adjusted life years*, a measure of overall disease burden, expressed as the number of years lost due to disease, disability or early death. The output above shows DALYs *saved* by the nudge, that is, the number of years *saved* due to *decreased* disease, disability and death.

As for the code: We basically told `nudge()` what kind of nudge we are envisioning, `"information"`, and what kind of behavior we are targeting: `"food"`-related behavior. We piped the result into `disease()` and told it what disease we are targeting: `"diabetes mellitus"`. Finally, `describe()` simplified the output to a brief description of the guesstimate, which is what you saw.

So, we got our guesstimate in terms of DALYs saved. There's many other measures we can get estimates for. Let's see how many *years lived with disability* or "YLDs" our nudge would save:

```{r}
nudge("information", behavior = "food") %>%
  disease("diabetes mellitus", measure = "YLDs") %>%
  describe
```

The reason why I picked YLDs is because it allows us to guesstimate the cost of healthcare saved by the nudge. We do that as follows, with `cost()`:

```{r}
nudge("information", behavior = "food") %>%
  disease("diabetes mellitus", measure = "YLDs") %>%
  cost %>%
  describe
```

Note that cost calculation is still a feature in development: for now it doesn't yet distinguish between the costs of different diseases. Run `?cost` to read more. 

Once we have the cost savings, we can also guesstimate the *net* savings in cost, that is, the savings from reduce healthcare costs minus the cost of the nudge, of running the campaign. We use the `net()` function:

```{r}
nudge("information", behavior = "food") %>%
  disease("diabetes mellitus", measure = "YLDs") %>%
  cost %>%
  net %>%
  describe
```

Next, say you are in charge of a campaign at an insurance company with 350,000 members. You can also calculate the net savings for 350,000 members, with the `per()` function:

```{r}
nudge("information", behavior = "food") %>%
  disease("diabetes mellitus", measure = "YLDs") %>%
  cost %>%
  net %>%
  per(350000) %>%
  describe
```

Finally, say you are considering running an experiment with your members, to see if your nudge has a positive impact. You can do quick power calculations with `sample_size()`. Here is how to do a power calculation for an experiment that wants to detect an impact on food-related behaviors.

```{r}
nudge("information", behavior = "food") %>%
  describe
nudge("information", behavior = "food") %>%
  sample_size
```

Here is how to do a power calculation for a much more ambitious experiment that wants to detect on impact on saving lives (preventing deaths):

```{r}
nudge("information", behavior = "food") %>%
  disease("diabetes mellitus", measure = "deaths") %>%
  describe
nudge("information", behavior = "food") %>%
  disease("diabetes mellitus", measure = "deaths") %>%
  sample_size
```

The low number of yearly deaths means you'll need a huge sample size! Better just look at behavior change!

Alright, that's a quick overview of what `better` can do for you!

<!-- END COPY -->
