---
output: github_document
---

<!-- Edit this to prevent git error: 9827634 -->
<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
options(
  better.gbd_path = "/Users/jwl38/Library/CloudStorage/GoogleDrive-janwillem.lindemans@gmail.com/My\ Drive/Offline\ Drive/R\ -\ Offline\ Drive/R\ packages\ JW/better/ignore/Global\ Burden\ of\ Disease\ Study\ Data"
  )
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# better

<!-- badges: start -->
<!-- badges: end -->

IMPORTANT: THIS PACKAGE IS STILL UNDER DEVELOPMENT. IT'S NOT YET INTENDED FOR USE. IT'S HERE FOR DEVELOPMENT PURPOSES ONLY.

The goal of better is to make the world a better place through impact evaluation.

## Installation

You can install the development version of better from [GitHub](https://github.com/) with:

``` r
# install.packages("devtools")
devtools::install_github("janlindemans/better")
```

## Set up

<!-- COPY from better vignette: edit original there -->

`better` has several functions that rely on data from the Global Burden of Disease study (GBD). This requires some setup.

### Download Global Burden of Disease codebook and data

Download the **GBD codebook** from the website of the Institute for Health Metrics and Evaluation (IHME) by clicking [here](
https://www.healthdata.org/sites/default/files/files/Data_viz/GBD_2019_Data_Tools_Guide.zip). 

Save the downloaded folder `IHME_GBD_2019_CODEBOOK` into a root folder in which you'll later also save GBD data. Let's call this root folder the *GBD folder*. You can create the GBD folder anywhere and name it whatever you like. Your GBD folder could be something like: `Users/johnwilliam/datasets/Global Burden of Disease`, with the codebook folder `Users/johnwilliam/datasets/Global Burden of Disease/IHME_GBD_2019_CODEBOOK`.

Download your first **GBD data** from the Global Health Data Exchange website. Click [here](https://vizhub.healthdata.org/gbd-results/) to go to the website. 

You'll need to create an account first. Click `Sign in`, then `Register`, etc.  

Then go back to the [GBD Results page](https://vizhub.healthdata.org/gbd-results/). It's a huge database, so you can't download everything. 

Start by downloading a basic dataset on *causes of death*, with the following parameters in the drop-down menus on the page:

* `GBD Estimate`: Keep that one as is, so `Cause of death or injury`.
* `Measure`: Keep `Deaths` and `DALYs`, and add `YLDs` from the drop-down.
* `Metric`: Fine as is, keep all three.
* `Cause`: Keep `All causes`, and add `Cardiovascular diseases` from the drop-down, under `Non-communicable diseases`, and `Diabetes and kidney diseases`.
* `Location`: Keep `Global`, and add `Low SDI`, `Middle SDI`, and `High SDI`, and `United States of America` (under `High-income North America`) and `Latin America and Caribbean` (below `Western Europe`).
* `Ages`: Keep `All ages`, and add `Age-standardized`, `<20 years`, `20+ years`, `<70 years` and `70+ years`.
* `Sex`: Keep `Both`, and add `Male` and `Female`.
* `Year`: Keep `2019`, and add `2009`.

Then click `Download`, and, for `Include ID Values`, you can keep `Both`. When it says "You can also monitor progress `here`", click `here` and then `Download`. 

Then go back by clicking the left arrow on top of the page, to the left of `Data Download`. It will take you back to the Search page, where you just entered all these parameters. It will remember all these parameters, so it's best you really go back like I described. 

Now download the same dataset, but with the following parameters:

* `Measure`: Select `YLLs`, `Prevalence`, and `Incidence`.

After you did that, go back and download a dataset on *risk factors*:

* `GBD Estimate`: Select `Risk factor`. 
* `Measure`: Keep `Deaths` and `DALYs`, and add `YLDs`. 
* `Metric`: Fine as is, keep all three.
* `Risk`: Keep `All risk factors`, add `behavioral risks`, `Tobacco` and `Dietary risks`.
* `Cause` until `Year`: Keep it all as is, if it remembered your selections from the dataset on causes of death (see above).

After you did that, go back, and download the same dataset but with `Measure` being `YLLs`.

After you did that, go back and download a dataset on *summary exposure values*:
* `GBD Estimate`: Select `Summary exposure values (SEV)`.
* For `Measure` and `Metric`, you won't have any choice.
* `Risk` until `Year`: Keep it all as is, if it remembered your selections from earlier on (see above).

After unzipping, the downloaded data folder should have a name similar to this: `IHME-GBD_2019_DATA-64c5a7b3-1`. Store this data folder anywhere in (a subfolder of) your GBD folder. For instance: `Users/johnwilliam/datasets/Global Burden of Disease/Raw GBD data/IHME-GBD_2019_DATA-64c5a7b3-1`. You'll do the same for future data downloads. Just don't change the names of the data files, and codebook files for that matter, because `better` searches for them using these names.

### Set global option in your `.Rprofile` file

Enter the path of your GBD folder (e.g., `Users/johnwilliam/datasets/Global Burden of Disease`) as the default GBD path via global options set in your `.Rprofile` file. It will be used by several `better` functions. It is not required, since you can also set global options manually in each session, or even set the path manually for each function call (see below). But it's easier to take care of this once and for all in `.Rprofile`. 

It will look like this:

```{r, eval = FALSE}
options(better.gbd_path = "/Users/johnwilliam/datasets/Global Burden of Disease")
```

If you haven't used `.Rprofile` before, it's a hidden file located somewhere like here: `Users/johnwilliam/.Rprofile`. Learn more about `.Rprofile` on [the Posit Support page](https://support.posit.co/hc/en-us/articles/360047157094-Managing-R-with-Rprofile-Renviron-Rprofile-site-Renviron-site-rsession-conf-and-repos-conf).

## Usage

<!-- COPY from better vignette: edit original there -->

The main purpose of `better` is to help you better guestimate the effect of health nudges, that is, behavioral interventions targeted at health behaviors. 

Imagine you are designing a public health campaign. You are thinking about focusing on preventing cardiovascular diseases, and you are considering making information about healthy lifestyle more accessible. What effects can you expect from a campaign like that? This is how we would do it with `better`.

First, we look up the effect of an information nudge on health behavior.

```{r}
library(tidyverse)
library(better)
nudge_to_behavior_d(nudge = "information", behavior = "health") 
```


This gives us a Cohen's *d*, a standardized effect size for the health information nudge, namely, *d* = `r nudge_to_behavior_d(nudge = "information", behavior = "health")$value`.

Second, as an intermediate step, we translate Cohen's *d* into a percentage point difference.

```{r}
nudge_to_behavior_d(nudge = "information", behavior = "health") %>%
  behavior_d_to_pct
```

This gives us a decrease of `r nudge_to_behavior_d(nudge = "information", behavior = "health") %>% behavior_d_to_pct %>% {round(abs(.$value),1)}` percentage points.

Finally, we can get our guestimate of the effect:

```{r}
nudge_to_behavior_d(nudge = "information", behavior = "health") %>%
  behavior_d_to_pct %>%
  behavior_pct_to_disease("cardiovascular diseases")
```

```{r}
gbd_cb <- gbd_codebook()
nudge_cb <- nudge_codebook()
nudge_cb$nudge
nudge_cb$behavior
```

We learned something! If we're going to roll out an information nudge in 100,000 people, targeted at cardiovascular diseases, we're expected to boost disability adjusted life years by about `r nudge_to_behavior_d(nudge = "information", behavior = "health") %>% behavior_d_to_pct %>% behavior_pct_to_disease("cardiovascular diseases") %>% {round(.$value)}` years! That's not bad. 

Note that this is based on the following statistic:

```{r}
gbd_describe(cause = "cardiovascular diseases")
```

Let's see, what if we would shift the focus of our intervention to diabetes and kidney diseases?

```{r}
nudge_to_behavior_d(nudge = "information", behavior = "health") %>%
  behavior_d_to_pct %>%
  behavior_pct_to_disease("diabetes and kidney diseases")
```

What if we tried a structural nudge?

```{r}
nudge_to_behavior_d(nudge = "structure", behavior = "health") %>%
  behavior_d_to_pct %>%
  behavior_pct_to_disease("diabetes and kidney diseases")
```
