---
title: "The Global Burden of Disease dataset"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{gbd}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
# Set all better options to NULL
for (i in stringr::str_subset(names(options()),"better\\.")) {
  print(i)
  print(.Options[[i]])
  .Options[[i]] <- NULL
  print(.Options[[i]])
}
options(
  rmarkdown.html_vignette.check_title = FALSE,
  better.gbd_path = NULL#"/Users/jwl38/Library/CloudStorage/GoogleDrive-janwillem.lindemans@gmail.com/My\ Drive/Offline\ Drive/R\ -\ Offline\ Drive/R\ packages\ JW/better/ignore/Global\ Burden\ of\ Disease\ Study\ Data"
  )
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
#devtools::build_rmd("vignettes/gbd.Rmd")
```

```{r, include = FALSE}
library(better)
library(tidyverse)
# So that the once-per-session messages are gone:
nudge("information") %>% 
  disease("cardiovascular diseases")
```

<!-- ORIGINAL -->

## Background

To make data-driven guesstimates, `better` uses the [Global Burden of Disease (GBD) dataset](https://www.healthdata.org/gbd/2019), an incredibly rich dataset on the impact of a long list of diseases. 

## Terms of use, license and citation

The GBD dataset is created and maintained by the [Institute for Health Metrics and Evaluation](https://www.healthdata.org/) at the University of Washington. The dataset is owned by the University of Washington. The `better` package uses the dataset in accordance with the [IHME free-of-charge non-commercial user agreement]( https://www.healthdata.org/Data-tools-practices/data-practices/ihme-free-charge-non-commercial-user-agreement).

Cite the data as follows: 

Global Burden of Disease Collaborative Network. Global Burden of Disease Study 2019 (GBD 2019) Results. Seattle, United States: Institute for Health Metrics and Evaluation (IHME), 2020. Available from [https://vizhub.healthdata.org/gbd-results/](https://vizhub.healthdata.org/gbd-results/).

<!-- end -->

## Quick intro to the Global Burden of Disease dataset

The GBD dataset contains data on death and disability, the diseases and injuries that cause them, and the risk factors for these causes.

```{r}
gbd
gbd_variables()
```

The `measure` variable refers to the measure of death and/or disability.

```{r}
unique(gbd$measure)
gbd_measures() %>%
  select(measure, explain)
```

The `metric` variable refers to the measure of death and/or disability.

```{r}
unique(gbd$metric)
gbd_measures() %>%
  select(measure, number, percent, rate)
```

```{r}
```


The `cause` variable refers to the cause of death or disability.

```{r}

unique(gbd$cause)
```


```{r}
```

```{r}
```

```{r}
```

```{r}
```

```{r}
```

```{r}
```

```{r}
```

```{r}
```

## Clean, save and load

`better` has functions that will clean the GBD data and codebook, and save it as a `gbd_dataset` object into an `.rds` file. You can then easily read-in this file.

```{r, eval = FALSE}
library(tidyverse) # for convenient tibble printing and more
library(better)
gbd_save()
gbd <- gbd_data()
gbd
```
```{r, echo = FALSE}
library(tidyverse)
library(better)
stopifnot(nchar(options()$better.gbd_path) > 0) # to
#gbd_save() # takes too long
warning()
gbd <- gbd_data()
gbd
stopifnot(nrow(gbd) > 0)
```

```{r, echo = FALSE}
# FOR TESTING
# stopifnot(nrow(gbd) > 0)
# get("gbd")
# t1 <- function(x) {
#   get("gbd")
# }
# t1()
# stopifnot(nrow(t1()) > 0)
# #stop(nrow(gbd))
# gbd_filter(cause = "cardiovascular diseases") # many columns...
```

Note I called the returned list `gbd`. It contains data (`.$gbd_data`) and the codebook (`.$gbd_codebook`). It will be the main input for several `better` functions, like the function discussed in the next section. 

## Get values from the GBD data

The core of the package is the `better::gbd_*` family of functions, revolving around the Global Burden of Disease dataset. The GBD *dataset* consists of the GBD *data* and the GBD *codebook*. In `better`, this dataset is available as the object `gbd`. `gbd` is a dataframe with a "codebook" attribute. `better` also gives the GBD dataset its own class, namely, `better_gbd`. The most important reason is that `better_gbd` objects are printed in a special way. 

`better::gbd_filter` gets values from the GBD data. 

By default, it uses `gbd` (which should be the `gbd_dataset` object) for its `data` argument.

```{r}
gbd_filter(cause = "cardiovascular diseases") # many columns...
```

Let's run it without specifying any of the parameters, sticking to the defaults. 

...

In case you don't know, `rei` stands for *risks, etiologies and impairments* in GBD-speak, and `val` is the value.

```{r}
gbd_filter(rei = "behavioral risks")
```

Yay, we learned something! What we found here is that behavioral risks cause a population of 100,000 to lose `r gbd_filter(rei = "behavioral risks") %>% pull(val) %>% round %>% format(big.mark=",")` Disability-Adjusted Life Years (DALYs) in one year! That's a lot of life years. 

```{r}
gbd_filter(cause = "cardiovascular diseases", rei = NA, measure = c("deaths", "DALYs"))
gbd_filter(cause = "cardiovascular diseases", rei = c("behavioral risks", "all risk factors"))
```

## Doing things differently

If so far so good, you can ignore this section. It's about doing things a bit differently - but I don't recommend that.

If you haven't set global options in `.Rprofile`, or in your current R session, you can still manually enter the path to your GBD folder.

```{r, eval = FALSE}
gbd_manual <- gbd_data("/Users/johnwilliam/datasets/Global Burden of Disease")
```
```{r, echo = FALSE, message = FALSE}
path_gbd <- paste0(
  "/Users/jwl38/Library/CloudStorage/Box-Box/CAH/CAH Shared/IRB Projects",
  "/Health Projects/Health Team/Public datasets on health",
  "/Global Burden of Disease Study Data"
)
gbd_manual <- gbd_data(path_gbd)
```

The result is the same:

```{r}
gbd_filter(data = gbd_manual) %>% 
  select(cause:measure)
```

The same goes for `better::gbd_save()`.

## Try

```{r}
library(better)
```

By default, `gbd_filter()` sets the `location` parameter to `"Global"`. 

```{r}
gbd_filter("diabetes mellitus")
gbd_filter("diabetes mellitus")$location
```

```{r}
options(
  better.location = "United States of America"
)
gbd_filter("diabetes mellitus")
gbd_filter("diabetes mellitus")$location
```

## Beyond the starter dataset

To learn more about how `better` uses GBD data in it's estimates, run `better_vignette(gbd)` to access the pre-installed HTML vignette. 

```{r, eval = FALSE}
library(better)
?better_vignette
better_vignette("dataset-extension")
```

Why not just have all of that info in this regular vignette here? Well, the vignette requires some setup (e.g., downloading extra GBD data). But regular vignettes are built from R code at the time of installation, and thus require everything to already be set up correctly. A pre-installed HTML is more forgiving. 

